Neutrino Torque

### 2018-10-09 neu_torque.tex
added figure

The spin of a halo in Lagrangian space $\bs{j}_L\propto\int_{V_L}\bs{q}\times\bs{\nabla}\phi\diff^3\bs{q}$ can be written in first order as $\bs{j}_I\propto\bs{\epsilon}\,\bold{I}\,\bold{T}$ \citep{1984ApJ...286...38W}, where $\phi$ is the gravitational potential, $\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$, $\bold{I}=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the inertia tensor, $\bold{T}=(T_{ij})=(\partial_i\partial_j\phi)$ is the tidal shear tensor, and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol. This picture shows that the initial spin of the protohalo is generated by the misalignment between the protogalactic inertia tensor and local gravitational shear tensor, and grows at first order in the early Universe \citep{1984ApJ...286...38W}. Actually, as halos collapse, nonlinear gravity field is hard to torque small objects, and we expect the spin directions are contributed mostly in the linear regime. These can be straightforwardly tested in numerical simulations. The halo merging and baryonic feedback, usually been discussed in the halo/galaxy alignment studies, do not affect the angular momentum conservation, because connected/disconnected matter and baryonic matter in Lagrangian space are torqued by the same gravitational shear.

@@

\tcb{\textit{Simulation.---}}
These correlations and coefficients are tested across a set of high-resolution $N$-body simulations \citep{2018ApJS..237...24Y}. Given any halo produced by the simulation, all the belonging particles are mapped back to Lagrangian space. The status of this definite set of particles can be traced in a resimulation, for example, their spin $\bs{j}(z)=\sum\bs{x}(z)\times\bs{v}(z)$ relative to their center of mass. In Fig.\ref{fig.1} we plot (note that all quantities are projected onto the plane of this letter) the LSS from a simulation at redshift $z=0$, and mark one selected halo with its $r_{200}$ radius (within which the mean halo density is 200 times the mean matter density of the universe) and its spin direction $\bs{j}(z=0)$ (all the spin arrows are normalized to have 15 Mpc$/h$). 
We illustrate Lagrangian mapping of this halo with its protohalo's column density (the orange clouds), with its initial spin direction $\bs{j}_L$. We can see that the Lagrangian spin $\bs{j}_L$ and Eulerian spin $\bs{j}_0$ have similar directions. \tcb{Indeed, by averaging over all the halos in the simulation, we find the expectation value of the cross-correlation coefficient $\left\langle \bs{j}(z) \cdot \bs{j}_L \right\rangle$ is smoothly decreasing from unity to 0.80.} This shows that the halo spin direction is conserved from the initial state all the way to the final state of the Universe.

>>>>

The spin of a halo in Lagrangian space $\bs{j}_L\propto\int_{V_L}\bs{q}\times-\bs{\nabla}\phi\diff^3\bs{q}$ can be written in first order as $\bs{j}_I\propto\bs{\epsilon}\,\bold{I}\,\bold{T}$ \citep{1984ApJ...286...38W}, where $\phi$ is the gravitational potential, $\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$, $\bold{I}=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the inertia tensor, $\bold{T}=(T_{ij})=(\partial_i\partial_j\phi)$ is the tidal shear tensor, and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol. This picture shows that the initial spin of the protohalo is generated by the misalignment between the protogalactic inertia tensor and local gravitational shear tensor, and grows at first order in the early Universe \citep{1984ApJ...286...38W}. Actually, as halos collapse, nonlinear gravity field is hard to torque small objects, and we expect the spin directions are contributed mostly in the linear regime. These can be straightforwardly tested in numerical simulations. The halo merging and baryonic feedback, usually been discussed in the halo/galaxy alignment studies, do not affect the angular momentum conservation, because connected/disconnected matter and baryonic matter in Lagrangian space are torqued by the same gravitational shear.

@@

\tcb{\textit{Simulation.---}}
These correlations and coefficients are tested across a set of high-resolution $N$-body simulations \citep{2018ApJS..237...24Y}. Given any halo produced by the simulation, all the belonging particles are mapped back to Lagrangian space. The status of this definite set of particles can be traced in a resimulation, for example, their spin $\bs{j}(z)=m_p\sum\bs{x}(z)\times\bs{v}(z)$ relative to their center of mass. In Fig.\ref{fig.1} we plot (note that all quantities are projected onto the plane of this letter) the LSS from a simulation at redshift $z=0$, and mark one selected halo with its $r_{200}$ radius (within which the mean halo density is 200 times the mean matter density of the universe) and its spin direction $\bs{j}(z=0)$ (all the spin arrows are normalized to have 15 Mpc$/h$). 
We illustrate Lagrangian mapping of this halo with its protohalo's column density (the orange clouds), with its initial spin direction $\bs{j}_L$. We can see that the Lagrangian spin $\bs{j}_L$ and Eulerian spin $\bs{j}_0$ have similar directions. \tcb{Indeed, by averaging over all the halos in the simulation, we find the expectation value of the cross-correlation coefficient $\left\langle \bs{j}(z) \cdot \bs{j}_L \right\rangle$ is smoothly decreasing from unity to 0.80.} This shows that the halo spin direction is conserved from the initial state all the way to the final state of the Universe.
